---
title: "TROPIRES Summer School Uganda August 19–23 2024"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Data for soil near-infrared calibration modeling exercises

Leonardo Ramirez-Lopez, Moritz Mainka, Laura Summerauer


# Data overview

## TropSOC data from Uganda (cropland and forest)

- Forest samples come from the Kibale National Park
- Cropland samples were taken from the Rwenzori foothills, West of Kibale National Park

```{r}
tropsoc_data <- readRDS("data/calibration_data/NIRabs_TropSOC_Uganda.rds")

str(tropsoc_data)
```

### Absorbance spectra in NIR range

- Units are wavenumbers (1 / cm)

```{r}
range(as.numeric(colnames(tropsoc_data$abs)))
```


## Samples from PhD project Laura

cropland samples only from:

- Rwenzori foothills (West of Kibale NP) 
- East of Kibale NP
- Saaka / volcanic area in Fort Portal (Saaka University campus)
- I very much recommend to **remove Saaka samples from modeling**, since these are substantially different from all the other samples!! 

```{r}
laura_data <- readRDS("data/calibration_data/NIRabs_PhDLaura_Uganda.rds")

str(laura_data)
```

# Calibration modeling

- Use here merged TropSOC and Laura samples (except volcanic Saaka samples)
- 151 samples

```{r}
alldata <- readRDS("data/calibration_data/NIRabs_allUganda_noSaaka.rds")

str(alldata)
```

## Pre-process / smoothen spectra

```{r}
wavs <- as.numeric(colnames(alldata$abs))
wavs_pre <- as.numeric(colnames(alldata$abs_pre))

alldata$abs_pre <- alldata$abs |>
  prospectr::movav(w = 19)
  # detrend(wav = wavs_pre, p = 2)
  # standardNormalVariate() 
  # savitzkyGolay(m = 2, p = 2, w = 17) 
```

## Visualization

```{r}
# raw absorbance
matplot(x = as.numeric(colnames(alldata$abs)), y = alldata$abs[1,],
        xlab = expression(paste("Wavenumber ", cm^{-1})),
        ylab = 'Absorbance',
        type = 'l',
        lty = 1, 
        main = "Raw spectra",
        # ylim = c(0.2, 0.8),
        xlim = c(7500, 3900))

# pre-processed absorbance data
matplot(x = as.numeric(colnames(alldata$abs_pre)), y = alldata$abs_pre[1,],
        xlab = expression(paste("Wavenumber ", cm^{-1})),
        ylab = 'Absorbance',
        type = 'l',
        lty = 1, 
        main = "Pre-processed spectra",
        # ylim = c(0.2, 0.8),
        xlim = c(7500, 3900))
```


## PLS modeling

```{r}
library(caret)
library(tidyverse)

# maxiumum number of components
pls_ncomp_max <- 10

## define trainControl
train_control <- trainControl(
  method = "repeatedcv",
  savePredictions = TRUE, selectionFunction = "oneSE"
)

alldata$rowIndex <- c(1:nrow(alldata))


## train a pls regression model
pls_model <- train(x = alldata$abs_pre,
                       y = alldata$TC_gkg,
                       method = "pls",
                       tuneLength = pls_ncomp_max,
                       trControl = train_control,
                       preProcess = c("center", "scale"))


pls_model
plot(pls_model)
range(alldata$TC_gkg)

```


```{r echo = FALSE}

source("R/pls_eval.R")

eval <- pls_val_plot(pls_object = pls_model, data = alldata, ID = "sample_id", property = "TC", unit = "g/kg", label = "SOC", val_type = "CV")
eval$p_calval
```

